Integrated approach to modeling to krill population dynamics in Wester Antarctic Peninsula: spatial and ecosystem considerations

### 2024-05-31 index.Rmd
adding html but not in page

```{r setup1, echo=FALSE}
rm(list = ls())
set.seed(999)
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center',
                      fig.pos = "H",
                      dev = 'jpeg',
                      dpi = 300,
                      tidy.opts=list(width.cutoff=50), 
                      tidy=TRUE)
#XQuartz is a mess, put this in your onload to default to cairo instead
options(bitmapType = "cairo") 
# (https://github.com/tidyverse/ggplot2/issues/2655)
# Lo mapas se hacen mas rapido
```

```{r message=FALSE, eval=TRUE}
# install.packages("devtools")
# devtools::install_github("r4ss/r4ss", ref="development")
# install.packages("caTools")
# library("caTools")
# install.packages("r4ss")
library(r4ss)
library(here)
#remotes::install_github("PIFSCstockassessments/ss3diags")
library(ss3diags)
library(kableExtra)
library(doParallel) # facilita la ejecución paralela en R
detectCores()
registerDoParallel(8)
library(ggpubr)
library(tibble)
library(openxlsx)
library(ggthemes)
library(forecast)
```

```{r message=FALSE, warning=FALSE}
dir01 <- here("s01") # agreggate data (no spatial diferences)
dir1<-here("s1") # Data strata flishery
dir2<-here("s2") # Same 9 with areas (SubStrata) as fleet. Dif size comoposition and dif CPUE and dif survey length and biomass data by strata
dir3<-here("s3") # whitout S-R
dir4<-here("s4") # 
dir5<-here("s5") # whitout S-R
dir6<-here("s6") # whitout S-R
dir7<-here("s7") # whitout S-R
dir8<-here("s8") # S2 set parametres EMM-2024/23 (Mardones)
```

@@

```
## Scenarios

@@

| Scenario | Description                                      |
|:---------:|:-------------------------------------------------|
|    s01     |Fishery and Survey (AMLR) data, Predator, Environmental aggregate data in 48.1 |
|    s1     | Fishery data (Length, Index, Catch) by strata  and Predator and Env data      |
|    s2    | Fishery and Survey (AMLR) data Length, Index, Catch by strata. Predator and Env data |
|    s3     | "s2" without S-R relation         |
|    s4     | "s2" Ricker S-R relation estimated          |
|    s5     | "s2" BH S-R relation weak (0.9 steepness)     |
|    s6     | "s2" BH S-R relation strong (0.6 steepness)     |
|    s7     | "s2" BH S-R relation mid-strong (0.65 steepness) estimated     |
|    s8     | "s2"  w/ set of parameters estimated in @EMM-204/32 |

@@

```{r}
SSplotSelex(base.model2,
            subplots = 1)
```

```{r eval=FALSE}
SSplotBiology(base.model2, 
              subplots =2, 
              labels = c("Length (cm)",
                         "Age (yr)",
                         "Maturity", 
                         "Mean weight (kg) in last year",
                          "Spawning output",
                         "Length (cm, beginning of the year)", 
                         "Natural mortality",
                          "Female weight (kg)", 
                         "Female length (cm)", 
                         "Fecundity", 
                         "Default fecundity label",
                          "Year", 
                         "Hermaphroditism transition rate", 
                         "Fraction females by age at equilibrium"),
 )

@@

aporte de las cohortes por año para las capturas.
```{r eval=FALSE}
SSplotCohortCatch(base.model2, subplots = 1)
```

\quad

AJuste de tallas por flota
```{r eval=FALSE}
SSplotComps(base.model2, subplots = 1)
```

Otros plots
```{r}
SSplotDynamicB0(base.model2, uncertainty = F)
#SSplotSPR(base.model3)
```

```{r eval=FALSE}
SSplotPars(base.model2)

Salida de las biomasas con las dos flotas

```{r}
SSplotTimeseries(base.model2, subplot = 7)
```
Salida de las biomasas con todas las flotas

```{r}
SSplotTimeseries(base.model2, subplot = 7)
```

@@

## Residual

```{r eval=FALSE}
SSplotRunstest(base.model8, 
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model2, 
               subplots = "len",
               add=T)
```

@@

## Predator fleet with RW

Random Walk (RW) refers to a mathematical model that describes a stochastic process in which a variable changes randomly over time, without a clear trend or pattern.

Specifically, a random walk can be used as a Bayesian estimation technique to infer the posterior distribution of an unknown parameter. In this approach, it is assumed that the prior distribution of the parameter is a normal distribution with mean zero and a known variance, and that the parameter value at each time step follows a random walk process. Based on the observed data and the prior distribution, the posterior distribution of the parameter can be calculated using Bayesian inference.

Random walk is a useful tool for parameter estimation in dynamic models, as it allows for modeling uncertainty and variability in the parameter's evolution over time. However, it is important to note that the random walk assumes that the changes in the parameter are random and without a clear trend, which may not be appropriate in all cases.

@@

```{r eval=FALSE}
retro(dir=dir7, oldsubdir="", 
      newsubdir="Retrospective", 
      years= 0:-5,
      exe="ss_osx",
      extras = "-nox", 
      skipfinished = F)
```

```{r echo=FALSE}
#s01
retroModels01 <- SSgetoutput(dirvec=file.path(dir01,
                                            "Retrospective",
                                            paste("retro",0:-5,
                                                  sep="")))

@@

@@

retro7 <- SSplotRetro(retroSummary7,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)

@@

```{r echo=FALSE}
ggarrange(retro01,
          retro1,
          retro2,
          retro3,
          retro4,
          retro5,
          retro6,
          retro7, ncol = 4, nrow = 4)
```

@@

Implementing the Hindcast Cross-Validation (HCxval) diagnostic in Stock Synthesis requires the same model outputs generated by `r4ss:SS_doRetro()`. As a robust measure of prediction skill, we implemented the mean absolute scaled error (MASE). In brief, the MASE score scales the mean absolute. Regarding (A MASE score > 1 indicates that the average model forecasts are worse than a random walk. Conversely, a MASE score of 0.5 indicates that the model forecasts twice as
accurately as a naïve baseline prediction; thus, the model has prediction skill.

```{r}
hci = SSplotHCxval(retroSummary2, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
```
## Kobe (status)

```{r}
mvln = SSdeltaMVLN(base.model8,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=1000)

mvln = SSdeltaMVLN(base.model2,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=1000)
```

@@

R0.vec <- seq(15,20,0.5)  

@@

Baseval <- round(Base$parameters$Value[grep("R0",Base$parameters$Label)],2)
Baselab <- paste(Baseval,sep="")
axis(1,at=Baseval,label=Baselab)
abline(v = Baseval, lty=2)
dev.off()

### SSplotComparisons() ----
# Comparación de series de tiempo 
labs <- paste("SR_Ln(R0) = ",R0.vec)
labs[which(round(R0.vec,2)==Baseval)] <- paste("SR_Ln(R0) = ",
                                               Baseval,"(Base model)")

@@

SSplotComparisons(summaryoutput,
                  legendlabels = c("No Spatial",
                                   "s1",
                                   "s2",
                 "s3",
                 "s4",
                 "s5",
                 "s6",
                 "s7",
                 "s8"),
                 labels = c("Year", 
                            "Spawning biomass (t)",
                            "Relative spawning biomass", 
                            "Age-0 recruits (1,000s)",
                            "Recruitment deviations", 
                            "Index", "Log index", 
                            "1 - SPR", 
                            "Density",
                            "Management target", 
                            "Minimum stock size threshold",
                            "Spawning output",
                            "Harvest rate"),
                 plotdir = "~/Figs")

@@

```
comparision between select models

@@

                 "s2",
                 "s3"),
                 labels = c("Year", 
                            "Spawning biomass (t)",
                            "Relative spawning biomass", 
                            "Age-0 recruits (1,000s)",
                            "Recruitment deviations", 
                            "Index", "Log index", 
                            "1 - SPR", 
                            "Density",
                            "Management target", 
                            "Minimum stock size threshold",
                            "Spawning output",
                            "Harvest rate"),
                 plotdir = "~/Figs")

@@

                 "s2",
                 "s3"),
                 subplots = c(2, 16),
                 plotdir = "~/Figs")
comp2_8
```
## Comparsion in sd long term time series

@@

ggplot(err_long, aes(x = variable, y = value)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, alpha = 0.3) +
  labs(title = "",
       x = "",
       y = "Spawning B.(mt)") +
  theme_few() +
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1))
base.model2$

@@

ggarrange(p01,p1,
          p2,
          p3,
          p4,
          p5,
          p6,
          p7,
          p8, ncol=3, nrow = 3,
          common.legend = T)
```

\newpage
# REFERENCES

>>>>

```{r setup1, echo=FALSE}
#rm(list = ls())
#set.seed(999)
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center',
                      fig.pos = "H",
                      dev = 'jpeg',
                      dpi = 300,
                      tidy.opts=list(width.cutoff=50), 
                      tidy=TRUE)
#XQuartz is a mess, put this in your onload to default to cairo instead
options(bitmapType = "cairo") 
# (https://github.com/tidyverse/ggplot2/issues/2655)
# Lo mapas se hacen mas rapido
```

```{r message=FALSE, eval=TRUE}
# install.packages("devtools")
# devtools::install_github("r4ss/r4ss", ref="development")
# install.packages("caTools")
# library("caTools")
# install.packages("r4ss")
library(r4ss)
library(here)
#remotes::install_github("PIFSCstockassessments/ss3diags")
library(ss3diags)
library(kableExtra)
library(doParallel) # facilita la ejecución paralela en R
detectCores()
registerDoParallel(8)
library(ggpubr)
library(tibble)
library(openxlsx)
library(ggthemes)
library(forecast)
library(tidyr)
library(mixR)
```

```{r message=FALSE, warning=FALSE}
dir01 <- here("s01") # agreggate data (no spatial diferences)
dir1<-here("s1") # Data strata flishery
dir2<-here("s2") # Same 9 with areas (SubStrata) as fleet. Dif size comoposition and dif CPUE and dif survey length and biomass data by strata
dir3<-here("s3") # whitout S-R
dir4<-here("s4") # 
dir5<-here("s5") # whitout S-R
dir6<-here("s6") # whitout S-R
dir7<-here("s7") # whitout S-R
dir8<-here("s8")
Figs <- here("Figs")# S2 set parametres EMM-2024/23 (Mardones)
```

@@

## Scenarios

@@

| Scenario | Description                                                                          |
|:------------:|:---------------------------------------------------------|
|   s01    | Fishery and Survey (AMLR) data, Predator, Environmental aggregate data in 48.1       |
|    s1    | Fishery data (Length, Index, Catch) by strata and Predator and Env data              |
|    s2    | Fishery and Survey (AMLR) data Length, Index, Catch by strata. Predator and Env data |
|    s3    | "s2" without S-R relation                                                            |
|    s4    | "s2" Ricker S-R relation estimated                                                   |
|    s5    | "s2" BH S-R relation weak (0.9 steepness)                                            |
|    s6    | "s2" BH S-R relation strong (0.6 steepness)                                          |
|    s7    | "s2" BH S-R relation mid-strong (0.65 steepness) estimated                           |
|    s8    | "s2" w/ set of parameters estimated in @EMM-204/32                                   |

@@

```{r}
SSplotSelex(base.model2,
            subplots = 1)
```
```{r}
SSplotComps(base.model01)
SSplotComps(base.model1)
SSplotComps(base.model2)
SSplotComps(base.model3)
SSplotComps(base.model4)
SSplotComps(base.model5)
SSplotComps(base.model6)
SSplotComps(base.model7)
SSplotComps(base.model8)

@@

@@

## Residual consistency in `s2`

```{r eval=FALSE}
SSplotRunstest(base.model2, 
               subplots = "cpue",
               add=T)
```

@@

## Relationship Stock-Recruit

```{r eval=FALSE}
l <- base.model4$recruit$SpawnBio
m <- base.model4$recruit$pred_recr
ggplot()+
  aes(l,m)+
  geom_point()+
  geom_smooth(method = "lm", 
              se =F)+
  ylim(0, max(base.model2$recruit$pred_recr))
```

@@

```{r eval=FALSE}
for (dir in directorios) {
  retro(
    dir = dir,
    oldsubdir = "",
    newsubdir = "Retrospective",
    years = 0:-5,
    exe = "ss_osx",
    extras = "-nox",
    skipfinished = FALSE
  )
}
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#s01
retroModels01 <- SSgetoutput(dirvec=file.path(dir01,
                                            "Retrospective",
                                            paste("retro",0:-5,
                                                  sep="")))

@@

#s8
retroModels8 <- SSgetoutput(dirvec=file.path(dir8,
                                            "Retrospective",
                                            paste("retro",0:-5,
                                                  sep="")))

retroSummary8 <- SSsummarize(retroModels8)

@@

retro7 <- SSplotRetro(retroSummary7,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)
retro8 <- SSplotRetro(retroSummary8,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)

@@

@@

Implementing the Hindcast Cross-Validation (HCxval) diagnostic in Stock
Synthesis requires the same model outputs generated by
`r4ss:SS_doRetro()`. As a robust measure of prediction skill, we
implemented the mean absolute scaled error (MASE). In brief, the MASE
score scales the mean absolute. Regarding (A MASE score \> 1 indicates
that the average model forecasts are worse than a random walk.
Conversely, a MASE score of 0.5 indicates that the model forecasts twice
as accurately as a naïve baseline prediction; thus, the model has
prediction skill.

```{r}
hci = SSplotHCxval(retroSummary2, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
```

## Kobe (status)

```{r}
mvln = SSdeltaMVLN(base.model01,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

mvln = SSdeltaMVLN(base.model1,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

mvln = SSdeltaMVLN(base.model2,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

mvln = SSdeltaMVLN(base.model3,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

mvln = SSdeltaMVLN(base.model4,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

mvln = SSdeltaMVLN(base.model5,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

mvln = SSdeltaMVLN(base.model6,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

mvln = SSdeltaMVLN(base.model7,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

mvln = SSdeltaMVLN(base.model8,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

@@

R0.vec <- seq(15,30,1)  

@@

Baseval <- round(Base$parameters$Value[grep("R0",Base$parameters$Label)],2)
Baselab <- paste(Baseval,sep="")
axis(1,at=Baseval,label=Baselab)
abline(v = Baseval, lty=2)
dev.off()
```

## SSplotComparisons()

# Comparación de series de tiempo 
labs <- paste("SR_Ln(R0) = ",R0.vec)
labs[which(round(R0.vec,2)==Baseval)] <- paste("SR_Ln(R0) = ",
                                               Baseval,"(Base model)")

@@

SSplotComparisons(summaryoutput,
                  legendlabels = c("No Spatial",
                                   "s1",
                                   "s2",
                 "s3",
                 "s4",
                 "s5",
                 "s6",
                 "s7",
                 "s8"),
                 labels = c("Year", 
                            "Spawning biomass (t)",
                            "Relative spawning biomass", 
                            "Age-0 recruits (1,000s)",
                            "Recruitment deviations", 
                            "Index", "Log index", 
                            "1 - SPR", 
                            "Density",
                            "Management target", 
                            "Minimum stock size threshold",
                            "Spawning output",
                            "Harvest rate"),
                 png=TRUE,plotdir=Figs)

@@

comparision between select models

@@

                 "s2",
                 "s3"),
                 labels = c("Year", 
                            "Spawning biomass (t)",
                            "Relative spawning biomass", 
                            "Age-0 recruits (1,000s)",
                            "Recruitment deviations", 
                            "Index", "Log index", 
                            "1 - SPR", 
                            "Density",
                            "Management target", 
                            "Minimum stock size threshold",
                            "Spawning output",
                            "Harvest rate"),
                 png=TRUE,plotdir=Figs)

@@

                 "s2",
                 "s3"),
                 subplots = c(2, 16))
comp2_8
```

## Comparsion in sd long term time series

@@

ggplot(err_long, aes(x = variable, y = value)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, alpha = 0.3) +
  labs(title = "",
       x = "",
       y = "Spawning B.(mt)") +
  theme_few() +
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1))
```

@@

ggarrange(p01,p1,
          p2,
          p3,
          p4,
          p5,
          p6,
          p7,
          p8, ncol=3, nrow = 3,
          common.legend = T)

## Recruit deviation

```{r}
dev01 <- base.model01$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S01")
dev1 <- base.model1$recruit[1:43,c(1,7)]%>% 
  mutate(Serie = "S1")
dev2 <- base.model2$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S2")
dev3 <- base.model3$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S3")
dev4 <- base.model4$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S4")
dev5 <- base.model5$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S5")
dev6 <- base.model6$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S6")
dev7 <- base.model7$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S7")
dev8 <- base.model8$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S8")

data_list <- list(dev01, dev1, dev2, dev3, dev4, dev5, dev6, dev7, dev8)
titles <- c("S01", "S1", "S2", "S3", "S4", "S5", "S6", "S7", "S8")
# Lista para almacenar los gráficos
plots <- list()

# Crear los gráficos en un ciclo for
for (i in 1:length(data_list)) {
  p <- ggplot(data_list[[i]], aes(x = Yr, y = dev)) +
    geom_point() +
    stat_smooth(method = "loess", 
                span = 0.15, 
                se = T) +
    geom_hline(yintercept = 0,
               col="red",
               linetype = "dotdash")+
    ggtitle(titles[i]) +
    theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1))+
    theme_few()
  plots[[i]] <- p
}

# Organizar los gráficos con ggarrange
combined_plot <- ggarrange(plotlist = plots, ncol = 3, nrow = 3)

combined_plot
```

```{r}
combined_data <- rbind(dev01, 
                           dev1, 
                           dev2, 
                           dev3, 
                           dev4,
                           dev5,
                           dev6,
                           dev7,
                           dev8)

combined_plot_join <- ggplot(combined_data, 
                             aes(x = Yr, 
                                 y = dev, 
                                 color = Serie)) +
  geom_point() +
  stat_smooth(method = "loess", 
              span = 0.15, 
              se = FALSE) +
  ggtitle("Combined Series Plot") +
  geom_hline(yintercept = 0,
               col="red",
               linetype = "dotdash")+
  scale_color_viridis_d(option="turbo",
                        name="Scenario")+
  theme_few() +
  theme(legend.position = "right")

combined_plot_join

\newpage

# REFERENCES
